---
title: "TS_Project"
author: "Deena Liz John"
date: "11/17/2017"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
library(ggplot2)
library(zoo)
library(reshape2)
library(corrplot)
library(car)
library(tseries)
library(forecast)
path <- "/Users/deena/Documents/Fall_Module2/TS_project/"
```

Reading in the data and converting it to TS object:

```{r}
df <- read.csv(file.path(path,"train.csv"), header = TRUE)

# Remove NAs and check consistency of number of rows
df <- na.omit(df)
nrow(df) == (2010-1986)*12

# Convert to TS object
BR_ts <- ts(df['Bankruptcy_Rate'], start = c(1987,1), end = c(2010, 12), frequency = 12)
```

#---------------------------------------------------------------------------------------------------------
# Exploratory Data Analysis (EDA)

Plotting Dependent variable: Bankruptcy Rate

```{r, fig.width=6, fig.height=3, warning=FALSE}
# Plotting TS data
ggplot() + geom_line(data=data.frame(BR_ts), aes(as.Date(as.yearmon(time(BR_ts))),BR_ts)) +
  labs(y="Bankruptcy Rate") + 
  labs(x="Month - Year") + 
  labs(title="Bankruptcy Rate across years")
```



Plotting all predictors across time:

```{r, fig.width=7, fig.height=6, warning=FALSE}
df2 <- df
df2[1] <- seq(as.Date("1987/01/01"), by = "month", length.out = 288)
df2_melt <- melt(df2, id.vars=c("Month"))
ggplot() + geom_line(data=df2_melt, aes(x=Month, y=value)) + facet_grid(variable ~ .,  scales = "free") + xlab("Month-Year") + ylab("Metric value")
```

The features `Population` and `House_Price_Index` have an upward trend and seem to be correlated with our predictor variable  `Bankruptcy_Rate`. On the other hand, `Unemployement_Rate` has no particular trend or correlation with `Bankruptcy_Rate`.

We can formally look at the correlation between features with a correlation matrix. 
Further, it is interesting to look at this matrix for different time frames for the purpose of prediction.


```{r}
# Entire timeframe: 1987 to 2010
M <- cor(df2[-1])
corrplot(M, method = "circle", tl.cex = 0.75, tl.col = "black")
```

```{r}
# Timeframe: 1995 to 2010
M <- cor(df2[df2['Month'] > '1995-01-01',][-1])
corrplot(M, method = "circle", tl.cex = 0.75, tl.col = "black")
```

```{r}
# Timeframe: 1993 to 2010
M <- cor(df2[df2['Month'] > '1993-01-01',][-1])
corrplot(M, method = "circle", tl.cex = 0.75, tl.col = "black")
```

For 1993 to 2010, `Bankruptcy_Rate` is much more correlated to the three features. Hence this time period could possibly be more suitable for a forecast model.

```{r}
par(mfrow=c(2,1))
acf(BR_ts, lag.max = 48, main = "ACF", ylab ="")
pacf(BR_ts, lag.max = 48, main = "PACF", ylab = "")
```

#---------------------------------------------------------------------------------------------------------#
# Feature Engineering

## Transforming Variables

```{r}
x <- df2[c('Population', 'House_Price_Index', 'Unemployment_Rate')]
xreg <- 100*(tail(x, -1) - head(x, -1))/head(x, -1)
xreg['Month'] <- df2['Month'][df2['Month'] > '1987-01-01']
```

## Split data into train and test:

Since we need to forecast for two years, let's go back two years and set that as the test data.

Additionally, setting two training datasets : One using the data from 1987, and the other from 1993.

Also, since the data is quite variable over time, let's take log of Bankruptcy Rate for the analysis.

```{r}
train_1987 <- window(log(BR_ts), start=c(1987,1), end=c(2008,12))
train_1993 <- window(log(BR_ts), start=c(1993,1), end=c(2008,12))

test <- window(log(BR_ts), start=c(2009,1), end=c(2010,12))
test2 <- window(BR_ts, start=c(2009,1), end=c(2010,12))
#train_1987 <- window(BR_ts, start=c(1987,1), end=c(2008,12))
#train_1993 <- window(BR_ts, start=c(1993,1), end=c(2008,12))

#test <- window(BR_ts, start=c(2009,1), end=c(2010,12))

```

# Model Building

Differencing the xreg variables in test:

```{r}
xreg_test_diff <- diff(as.matrix(df2[df2['Month'] > '2009-01-01' & df2['Month'] < '2011-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')]), lag = 12, differences = 1)
xreg_test <- df2[df2['Month'] >= '2009-01-01' & df2['Month'] < '2011-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')]

xreg_test_change_diff <- diff(as.matrix(xreg[xreg['Month'] > '2009-01-01' & xreg['Month'] < '2011-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')]), lag = 12, differences = 1)
xreg_test_change <- xreg[xreg['Month'] >= '2009-01-01' & xreg['Month'] < '2011-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')]
```


Fitting auto SARIMAX models on Training data from 1993:

No differencing:
```{r}
m1 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')])
m2 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index')])
m3 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index', 'Unemployment_Rate')])
m4 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population','Unemployment_Rate')])
m5 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population')])
m6 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index')])
m7 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate')])

pred1 <-  predict(m1, n.ahead = 24, newxreg = xreg_test)
pred2 <-  predict(m2, n.ahead = 24, newxreg = xreg_test[,c('Population', 'House_Price_Index')])
pred3 <-  predict(m3, n.ahead = 24, newxreg = xreg_test[,c('House_Price_Index', 'Unemployment_Rate')])
pred4 <-  predict(m4, n.ahead = 24, newxreg = xreg_test[,c('Population','Unemployment_Rate')])
pred5 <-  predict(m5, n.ahead = 24, newxreg = xreg_test[,c('Population')])
pred6 <-  predict(m6, n.ahead = 24, newxreg = xreg_test[,c('House_Price_Index')])
pred7 <-  predict(m7, n.ahead = 24, newxreg = xreg_test[,c('Unemployment_Rate')])

sqrt(mean((exp(pred1$pred) - exp(test))^2))
sqrt(mean((exp(pred2$pred) - exp(test))^2))
sqrt(mean((exp(pred3$pred) - exp(test))^2))
sqrt(mean((exp(pred4$pred) - exp(test))^2))
sqrt(mean((exp(pred5$pred) - exp(test))^2))
#sqrt(mean((exp(pred6$pred) - exp(test))^2))
sqrt(mean((exp(pred7$pred) - exp(test))^2))
```

With differencing d=1 and D=1:

```{r}
m1 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')], d=1, D=1)
m2 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index')], d=1, D=1)
m3 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index', 'Unemployment_Rate')], d=1, D=1)
m4 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population','Unemployment_Rate')], d=1, D=1)
m5 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population')], d=1, D=1)
m6 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index')], d=1, D=1)
m7 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate')], d=1, D=1)

pred1 <-  predict(m1, n.ahead = 24, newxreg = xreg_test)
pred2 <-  predict(m2, n.ahead = 24, newxreg = xreg_test[,c('Population', 'House_Price_Index')])
pred3 <-  predict(m3, n.ahead = 24, newxreg = xreg_test[,c('House_Price_Index', 'Unemployment_Rate')])
pred4 <-  predict(m4, n.ahead = 24, newxreg = xreg_test[,c('Population','Unemployment_Rate')])
pred5 <-  predict(m5, n.ahead = 24, newxreg = xreg_test[,c('Population')])
pred6 <-  predict(m6, n.ahead = 24, newxreg = xreg_test[,c('House_Price_Index')])
pred7 <-  predict(m7, n.ahead = 24, newxreg = xreg_test[,c('Unemployment_Rate')])

sqrt(mean((exp(pred1$pred) - exp(test))^2))
sqrt(mean((exp(pred2$pred) - exp(test))^2))
sqrt(mean((exp(pred3$pred) - exp(test))^2))
sqrt(mean((exp(pred4$pred) - exp(test))^2))
sqrt(mean((exp(pred5$pred) - exp(test))^2))
sqrt(mean((exp(pred6$pred) - exp(test))^2))
sqrt(mean((exp(pred7$pred) - exp(test))^2))
```

With differencing D=1:

```{r}
m1 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')], D=1)
m2 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index')], D=1)
m3 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index', 'Unemployment_Rate')], D=1)
m4 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population','Unemployment_Rate')], D=1)
m5 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population')], D=1)
m6 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index')], D=1)
m7 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate')],  D=1)

#pred1 <-  predict(m1, n.ahead = 24, newxreg = xreg_test)
#pred2 <-  predict(m2, n.ahead = 24, newxreg = xreg_test[,c('Population', 'House_Price_Index')])
#pred3 <-  predict(m3, n.ahead = 24, newxreg = xreg_test[,c('House_Price_Index', 'Unemployment_Rate')])
pred4 <-  predict(m4, n.ahead = 24, newxreg = xreg_test[,c('Population','Unemployment_Rate')])
pred5 <-  predict(m5, n.ahead = 24, newxreg = xreg_test[,c('Population')])
pred6 <-  predict(m6, n.ahead = 24, newxreg = xreg_test[,c('House_Price_Index')])
pred7 <-  predict(m7, n.ahead = 24, newxreg = xreg_test[,c('Unemployment_Rate')])

#sqrt(mean((exp(pred1$pred) - exp(test))^2))
#sqrt(mean((exp(pred2$pred) - exp(test))^2))
#sqrt(mean((exp(pred3$pred) - exp(test))^2))
sqrt(mean((exp(pred4$pred) - exp(test))^2))
sqrt(mean((exp(pred5$pred) - exp(test))^2))
sqrt(mean((exp(pred6$pred) - exp(test))^2))
#sqrt(mean((exp(pred7$pred) - exp(test))^2))
```


Training data from 1987:

```{r}
m1 <- auto.arima(train_1987, xreg = df2[df2['Month'] >= '1987-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')])
m2 <- auto.arima(train_1987, xreg = df2[df2['Month'] >= '1987-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index')])
m3 <- auto.arima(train_1987, xreg = df2[df2['Month'] >= '1987-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index', 'Unemployment_Rate')])
m4 <- auto.arima(train_1987, xreg = df2[df2['Month'] >= '1987-01-01' & df2['Month'] < '2009-01-01',][c('Population','Unemployment_Rate')])
m5 <- auto.arima(train_1987, xreg = df2[df2['Month'] >= '1987-01-01' & df2['Month'] < '2009-01-01',][c('Population')])
m6 <- auto.arima(train_1987, xreg = df2[df2['Month'] >= '1987-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index')])
m7 <- auto.arima(train_1987, xreg = df2[df2['Month'] >= '1987-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate')])

pred1 <-  predict(m1, n.ahead = 24, newxreg = xreg_test)
pred2 <-  predict(m2, n.ahead = 24, newxreg = xreg_test[,c('Population', 'House_Price_Index')])
pred3 <-  predict(m3, n.ahead = 24, newxreg = xreg_test[,c('House_Price_Index', 'Unemployment_Rate')])
pred4 <-  predict(m4, n.ahead = 24, newxreg = xreg_test[,c('Population','Unemployment_Rate')])
pred5 <-  predict(m5, n.ahead = 24, newxreg = xreg_test[,c('Population')])
pred6 <-  predict(m6, n.ahead = 24, newxreg = xreg_test[,c('House_Price_Index')])
pred7 <-  predict(m7, n.ahead = 24, newxreg = xreg_test[,c('Unemployment_Rate')])

sqrt(mean((exp(pred1$pred) - exp(test))^2))
sqrt(mean((exp(pred2$pred) - exp(test))^2))
sqrt(mean((exp(pred3$pred) - exp(test))^2))
sqrt(mean((exp(pred4$pred) - exp(test))^2))
sqrt(mean((exp(pred5$pred) - exp(test))^2))
#sqrt(mean((exp(pred6$pred) - exp(test))^2))
sqrt(mean((exp(pred7$pred) - exp(test))^2))
```



## Model Building with transformed variables (taken % change)

Fitting auto SARIMAX models on Training data from 1993:

No differencing:

```{r}
m1 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')])
m2 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population', 'House_Price_Index')])
m3 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('House_Price_Index', 'Unemployment_Rate')])
m4 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population','Unemployment_Rate')])
m5 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population')])
m6 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('House_Price_Index')])
m7 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Unemployment_Rate')])

pred1 <-  predict(m1, n.ahead = 24, newxreg = xreg_test_change)
pred2 <-  predict(m2, n.ahead = 24, newxreg = xreg_test_change[,c('Population', 'House_Price_Index')])
pred3 <-  predict(m3, n.ahead = 24, newxreg = xreg_test_change[,c('House_Price_Index', 'Unemployment_Rate')])
pred4 <-  predict(m4, n.ahead = 24, newxreg = xreg_test_change[,c('Population','Unemployment_Rate')])
pred5 <-  predict(m5, n.ahead = 24, newxreg = xreg_test_change[,c('Population')])
pred6 <-  predict(m6, n.ahead = 24, newxreg = xreg_test_change[,c('House_Price_Index')])
pred7 <-  predict(m7, n.ahead = 24, newxreg = xreg_test_change[,c('Unemployment_Rate')])

sqrt(mean((exp(pred1$pred) - exp(test))^2))
sqrt(mean((exp(pred2$pred) - exp(test))^2))
sqrt(mean((exp(pred3$pred) - exp(test))^2))
sqrt(mean((exp(pred4$pred) - exp(test))^2))
sqrt(mean((exp(pred5$pred) - exp(test))^2))
sqrt(mean((exp(pred6$pred) - exp(test))^2))
sqrt(mean((exp(pred7$pred) - exp(test))^2))
```

With differencing d=1 and D=1:

```{r}
m1 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')], d=1, D=1)
m2 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population', 'House_Price_Index')],d=1, D=1)
m3 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('House_Price_Index', 'Unemployment_Rate')], d=1, D=1)
m4 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population','Unemployment_Rate')], d=1, D=1)
m5 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population')],d=1, D=1)
m6 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('House_Price_Index')],d=1, D=1)
m7 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Unemployment_Rate')],d=1, D=1)

pred1 <-  predict(m1, n.ahead = 24, newxreg = xreg_test_change)
pred2 <-  predict(m2, n.ahead = 24, newxreg = xreg_test_change[,c('Population', 'House_Price_Index')])
pred3 <-  predict(m3, n.ahead = 24, newxreg = xreg_test_change[,c('House_Price_Index', 'Unemployment_Rate')])
pred4 <-  predict(m4, n.ahead = 24, newxreg = xreg_test_change[,c('Population','Unemployment_Rate')])
pred5 <-  predict(m5, n.ahead = 24, newxreg = xreg_test_change[,c('Population')])
pred6 <-  predict(m6, n.ahead = 24, newxreg = xreg_test_change[,c('House_Price_Index')])
pred7 <-  predict(m7, n.ahead = 24, newxreg = xreg_test_change[,c('Unemployment_Rate')])

sqrt(mean((exp(pred1$pred) - exp(test))^2))
sqrt(mean((exp(pred2$pred) - exp(test))^2))
sqrt(mean((exp(pred3$pred) - exp(test))^2))
sqrt(mean((exp(pred4$pred) - exp(test))^2))
sqrt(mean((exp(pred5$pred) - exp(test))^2))
sqrt(mean((exp(pred6$pred) - exp(test))^2))
sqrt(mean((exp(pred7$pred) - exp(test))^2))
```

With differencing D=1:

```{r}
m1 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')], D=1)
m2 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population', 'House_Price_Index')], D=1)
m3 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('House_Price_Index', 'Unemployment_Rate')], D=1)
m4 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population','Unemployment_Rate')], D=1)
m5 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population')], D=1)
m6 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('House_Price_Index')], D=1)
m7 <- auto.arima(train_1993, xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Unemployment_Rate')], D=1)

#pred1 <-  predict(m1, n.ahead = 24, newxreg = xreg_test_change)
#pred2 <-  predict(m2, n.ahead = 24, newxreg = xreg_test_change[,c('Population', 'House_Price_Index')])
#pred3 <-  predict(m3, n.ahead = 24, newxreg = xreg_test_change[,c('House_Price_Index', 'Unemployment_Rate')])
pred4 <-  predict(m4, n.ahead = 24, newxreg = xreg_test_change[,c('Population','Unemployment_Rate')])
pred5 <-  predict(m5, n.ahead = 24, newxreg = xreg_test_change[,c('Population')])
pred6 <-  predict(m6, n.ahead = 24, newxreg = xreg_test_change[,c('House_Price_Index')])
pred7 <-  predict(m7, n.ahead = 24, newxreg = xreg_test_change[,c('Unemployment_Rate')])

#sqrt(mean((exp(pred1$pred) - exp(test))^2))
#sqrt(mean((exp(pred2$pred) - exp(test))^2))
#sqrt(mean((exp(pred3$pred) - exp(test))^2))
sqrt(mean((exp(pred4$pred) - exp(test))^2))
sqrt(mean((exp(pred5$pred) - exp(test))^2))
sqrt(mean((exp(pred6$pred) - exp(test))^2))
sqrt(mean((exp(pred7$pred) - exp(test))^2))
```
0.003700048
0.003644708

Best models so far:

```{r}
#0.003323823
f1 <- auto.arima(train_1993, xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate')], d=1, D=1)
f1_pred <-  predict(f1, n.ahead = 24, newxreg = xreg_test[c('Unemployment_Rate')])
sqrt(mean((exp(f1_pred$pred) - exp(test))^2))
#ARIMA(2,1,1)(2,1,2)[12]

#NOTE: ALl models built with D=1, d=1 using % change on 1993 train data gave best and equally good results.
# All are ARIMA(1,1,1)(2,1,2)[12] models with different variable combinations used
#Best with only unemployement, followed by with each other variables individually

```

Plot fitted values:
```{r}
f <- exp(train_1993 - f1$residuals) #fitted values
par(mfrow=c(1,1))
plot(exp(train_1993), type = "l")
points(f, type = "l", col = "red")
legend("bottomright", legend = c("Observed", "Predicted"), lty = 1, col = c("black", "red"), cex = 0.5)
```

Plot whole data and prediction:
```{r}
par(mfrow=c(1,1))
plot(BR_ts, type = "l")
points(exp(f1_pred$pred), type = "l", col = "red")
```


##Trying models manually

Trials with normal data:

```{r}
#ARIMA(2,1,1)(2,1,2)[12] - 0.003323823 with only unemployement

m11 <- arima(train_1993, order = c(2,1,1), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')])
m11_pred <-  predict(m11, n.ahead = 24, newxreg = xreg_test[c('Population', 'House_Price_Index', 'Unemployment_Rate')])
sqrt(mean((exp(m11_pred$pred) - exp(test))^2))

m12 <- arima(train_1993, order = c(2,1,1), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate', 'Population')])
m12_pred <-  predict(m12, n.ahead = 24, newxreg = xreg_test[c('Unemployment_Rate', 'Population')])
sqrt(mean((exp(m12_pred$pred) - exp(test))^2))

m13 <- arima(train_1993, order = c(2,1,1), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate', 'House_Price_Index')])
m13_pred <-  predict(m13, n.ahead = 24, newxreg = xreg_test[c('Unemployment_Rate', 'House_Price_Index')])
sqrt(mean((exp(m13_pred$pred) - exp(test))^2))

m14 <- arima(train_1993, order = c(2,1,1), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population', 'House_Price_Index')])
m14_pred <-  predict(m14, n.ahead = 24, newxreg = xreg_test[c('Population', 'House_Price_Index')])
sqrt(mean((exp(m14_pred$pred) - exp(test))^2))

m15 <- arima(train_1993, order = c(2,1,1), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index')])
m15_pred <-  predict(m15, n.ahead = 24, newxreg = xreg_test[c('House_Price_Index')])
sqrt(mean((exp(m15_pred$pred) - exp(test))^2))

m16 <- arima(train_1993, order = c(2,1,1), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Population')])
m16_pred <-  predict(m16, n.ahead = 24, newxreg = xreg_test[c('Population')])
sqrt(mean((exp(m16_pred$pred) - exp(test))^2))

m17 <- arima(train_1993, order = c(1,1,1), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate')])
m17_pred <-  predict(m17, n.ahead = 24, newxreg = xreg_test[c('Unemployment_Rate')])
sqrt(mean((exp(m17_pred$pred) - exp(test))^2))
```

# Trail with % changed data:

```{r}
#ARIMA(1,1,1)(2,1,2)[12] -- 0.003586425 with only Unemployement
#ARIMA(2,1,1)(2,1,2)[12] -- 0.004143049
#ARIMA(2,1,2)(2,1,2)[12] -- 0.00421501


m11 <- arima(train_1993, order = c(2,1,2), seasonal = list(order = c(2,1,2), period = 12), xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population', 'House_Price_Index', 'Unemployment_Rate')])
m11_pred <-  predict(m11, n.ahead = 24, newxreg = xreg_test_change[c('Population', 'House_Price_Index', 'Unemployment_Rate')])
sqrt(mean((exp(m11_pred$pred) - exp(test))^2))

m12 <- arima(train_1993, order = c(2,1,2), seasonal = list(order = c(2,1,2), period = 12), xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Unemployment_Rate', 'Population')])
m12_pred <-  predict(m12, n.ahead = 24, newxreg = xreg_test_change[c('Unemployment_Rate', 'Population')])
sqrt(mean((exp(m12_pred$pred) - exp(test))^2))

m13 <- arima(train_1993, order = c(2,1,2), seasonal = list(order = c(2,1,2), period = 12), xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Unemployment_Rate', 'House_Price_Index')])
m13_pred <-  predict(m13, n.ahead = 24, newxreg = xreg_test_change[c('Unemployment_Rate', 'House_Price_Index')])
sqrt(mean((exp(m13_pred$pred) - exp(test))^2))

m14 <- arima(train_1993, order = c(2,1,2), seasonal = list(order = c(2,1,2), period = 12), xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population', 'House_Price_Index')])
m14_pred <-  predict(m14, n.ahead = 24, newxreg = xreg_test_change[c('Population', 'House_Price_Index')])
sqrt(mean((exp(m14_pred$pred) - exp(test))^2))

m15 <- arima(train_1993, order = c(2,1,2), seasonal = list(order = c(2,1,2), period = 12), xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('House_Price_Index')])
m15_pred <-  predict(m15, n.ahead = 24, newxreg = xreg_test_change[c('House_Price_Index')])
sqrt(mean((exp(m15_pred$pred) - exp(test))^2))

m16 <- arima(train_1993, order = c(2,1,2), seasonal = list(order = c(2,1,2), period = 12), xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Population')])
m16_pred <-  predict(m16, n.ahead = 24, newxreg = xreg_test_change[c('Population')])
sqrt(mean((exp(m16_pred$pred) - exp(test))^2))

m17 <- arima(train_1993, order = c(1,1,1), seasonal = list(order = c(2,1,2), period = 12), xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Unemployment_Rate')])
m17_pred <-  predict(m17, n.ahead = 24, newxreg = xreg_test_change[c('Unemployment_Rate')])
sqrt(mean((exp(m17_pred$pred) - exp(test))^2))
```


#Trying with only unemployement:

```{r}
#0.003323823 -- (2,1,1)(2,1,2)
#0.003306139 -- (2,1,2)(2,1,2)
#0.003266324 -- (0,1,1)(2,1,2)
#0.003219882 -- (1,1,1)(2,1,2)
#0.003172182 -- (0,1,2)(2,1,2)
#0.003163958 -- (0,1,3)(2,1,2)
#0.003137944 -- (0,1,4)(2,1,2)
#0.002599472 -- (0,1,15)(2,1,2)
#0.002575927 -- (0,1,14)(2,1,2)
#0.00253076 -- (0,1,13)(2,1,2)

m17 <- arima(train_1993, order = c(0,1,13), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate')])
m17_pred <-  predict(m17, n.ahead = 24, newxreg = xreg_test[c('Unemployment_Rate')])
sqrt(mean((exp(m17_pred$pred) - exp(test))^2))

#0.003586425 -- (1,1,1)(2,1,2)
#0.003345094 -- (1,1,3)(2,1,2)

#m17 <- arima(train_1993, order = c(0,1,5), seasonal = list(order = c(2,1,2), period = 12), xreg = xreg[xreg['Month'] >= '1993-01-01' & xreg['Month'] < '2009-01-01',][c('Unemployment_Rate')])
#m17_pred <-  predict(m17, n.ahead = 24, newxreg = xreg_test_change[c('Unemployment_Rate')])
#sqrt(mean((exp(m17_pred$pred) - exp(test))^2))

```

Best Model:
```{r}
m_final <- arima(train_1993, order = c(0,1,13), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1993-01-01' & df2['Month'] < '2009-01-01',][c('Unemployment_Rate')])
m_final_pred <-  predict(m_final, n.ahead = 24, newxreg = xreg_test[c('Unemployment_Rate')])
sqrt(mean((exp(m_final_pred$pred) - exp(test))^2))
```

```{r}
log(BR_ts)
test_df <- read.csv(file.path(path,"test.csv"), header = TRUE)

m_new <- arima(log(BR_ts), order = c(0,1,13), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[c('Unemployment_Rate')])

m_pred_2yrs <-  predict(m_new, n.ahead = 24, newxreg = test_df[c('Unemployment_Rate')])
sqrt(mean((exp(m_final_pred$pred) - exp(test))^2))

```

```{r}
m_final <- arima(exp(train_1987), order = c(1,1,1), seasonal = list(order = c(2,1,2), period = 12), xreg = df2[df2['Month'] >= '1987-01-01' & df2['Month'] < '2009-01-01',][c('House_Price_Index')])
m_final_pred <-  predict(m_final, n.ahead = 24, newxreg = xreg_test[c('House_Price_Index')])
sqrt(mean((exp(m_final_pred$pred) - exp(test))^2))
#sqrt(mean((m_final_pred$pred - exp(test))^2))
#0.003295595
#0.008353351
```

Plot best data and prediction:
```{r}
par(mfrow=c(1,1))
plot(BR_ts, type = "l", main = "RMSE 0.00253076 - SARIMA(0,1,13)(2,1,2)[12]")
points(exp(m_final_pred$pred), type = "l", col = "red")
```

## Checking model assumptions:

### (i) Test for Zero-Mean: t test

```{r, comment=''}
e <- exp(m_final$residuals)
#e <- m_final$residuals
t.test(e)
```

Clearly p-value is high considering a significance level of 0.05. Hence we cannot reject the null hypothesis and the residuals have zero mean. This assumption is satisfied.

### (ii) Homoscedasticity

```{r, comment=''}
par(mfrow=c(1,1))
plot(e, main="Residuals vs t", ylab="")
#abline(v=c(1966, 1976, 1986), lwd=3, col="red")
group <- c(rep(1,40),rep(2,40),rep(3,40),rep(4,40), rep(5,32))
levene.test(exp(e), group)#Levene
```

### (iii) Zero-Correlation

```{r, comment=''}
tsdiag(m_final)
```

### (iv) Normality

```{r, comment=''}
par(mfrow=c(1,1))
qqnorm(e, main="QQ-plot of Residuals")
qqline(e, col = "red")
shapiro.test(e) #SW test
```


Also, the p-value from Shapiro-Wilk test is greater than 0.05 (assumed significance level) and so we fail to reject the null hypothesis. Hence the normality assumption is satisfied.

